### 2018-12-08 paper.tex
Add latex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{multicol}
\setlength{\columnsep}{1cm}

\begin{document}
    \begin{center}
        \textbf{Evaluation of Classifiers on Spam Filtering} 
    \end{center}

    \begin{center}
        Griffin Dunn and William Thompson
    \end{center}

    \textbf{Abstract:} Spam Filters rely on a variety of techniques to work effectively.

    \begin{multicols}{2}
        \section{Introduction}
    \end{multicols}

\end{document}

### 2018-12-09 paper/paper.tex
Move latex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{sectsty}
\usepackage{multicol}
\setlength{\columnsep}{1cm}
\sectionfont{\fontsize{12}{15}\selectfont}

\begin{document}
    \begin{center}
        \textbf{Evaluation of Classifiers and Feature Selection on Spam Filtering} 
    \end{center}

    \begin{center}
        Griffin Dunn and William Thompson
    \end{center}

    \textbf{Abstract:} 
        In this paper we explore the effects of extracting features that are not solely based on word frequency.

    \begin{multicols}{2}
        \section{Introduction}

        \section{Dataset Selection}

        \section{Data Preparation}

        \section{Feature Extraction}

        \section{Classifier Selection}

        \section{Classifier Tuning}

        \section{Results}

        \section{Evaluation}

        \section{Conclusion}

        \section{References}

    \end{multicols}

\end{document}

### 2018-12-09 paper/paper.tex
Added bib, started how to run and added some info on data preperation

@@
 \usepackage{multicol}
+\usepackage{apacite}
 \setlength{\columnsep}{1cm}
+\usepackage[margin=0.5in]{geometry} % don't know what size margins to use
+
 \sectionfont{\fontsize{12}{15}\selectfont}
@@
         \section{Dataset Selection}
+            The correct selection of a dataset was crucial to the testing of 
+            different types of features. There were many characteristics that were
+            on our "wishlist." Such as having the freedom of processing 
+            the raw text ourselves and determining what features may be important, 
+            and not have the data preproccesed or turned into vectors.
+            Next we wanted data that was conversational, that didn't use too much 
+            slang or leetspeak and was in general proper English. The size of the
+            dataset needed to be fairly large, at least 1000, for frequency analysis.
+             Lastly, we would like a dataset that was referenced in multiple papers to make comparisons
+            of our results to others. So we shortlisted a few datasets that fulfilled
+            some of these characteristics.
 
+            The UCI SMS Spam Dataset is a collection of conversational text messages
+            origninating from a variety of sources.
         \section{Data Preparation}
-
+            Once data was loaded in we started by running the data the data through
+            the Python Standard Library's mail parser, which striped the mail headers
+            and put them into a map for separate analysis. It also handled multipart
+            data as files were encoded into the file in base64. Without using a mail
+            parser it would be difficult to just get the body of the text which
+            would be used for feature extraction. Other papers left this data in the
+            email, possibly because this information is hard to extract and the tools
+            weren't as easily accessable when their implementation was written. We
+            decided to strip this data out as to not add noise and also because many 
+            headers were left on from other spam checkers, identifying whether or not
+            it believed the message was spam or ham.
         \section{Feature Extraction}
@@
     \end{multicols}
-
+    %\bibliographystyle{apacite}
+    %\bibliography{spam.bib}
 \end{document}

### 2018-12-09 paper/paper.tex
Added process image, added information about Data Preperation and Intro

@@
 \documentclass[12pt]{article}
+
 \usepackage[utf8]{inputenc}
 \usepackage[english]{babel}
-
 \usepackage{sectsty}
-\usepackage{multicol}
-\usepackage{apacite}
-\setlength{\columnsep}{1cm}
+\usepackage{multicol, caption}
+\usepackage{graphicx}
+\usepackage{wrapfig}
+%\usepackage{biblatex}
 \usepackage[margin=0.5in]{geometry} % don't know what size margins to use
 
+\newenvironment{Figure}
+  {\par\medskip\noindent\minipage{\linewidth}}
+  {\endminipage\par\medskip}
+
+\setlength{\columnsep}{1cm}
+%\addbibresource{paper.bib}
+%\nocite{*}
 \sectionfont{\fontsize{12}{15}\selectfont}
@@
     \begin{center}
-        \textbf{Evaluation of Classifiers and Feature Selection on Spam Filtering} 
+        \textbf{Evaluation of Classifiers and Feature Selection on Spam Filtering test} 
     \end{center}
@@
     \textbf{Abstract:} 
-        In this paper we explore the effects of extracting features that are not solely based on word frequency.
+        In this paper we explore the effects of extracting features that are not 
+        solely based on word frequency and the selection of classifier for the task
+        of determining if a message is spam or not spam (ham). These methods
+        will be tested on a publicly available corpora and compared to the results
+        of other models introduced by other authors. Classifiers will be selected
+        from the variety available through the open-source Python module scikit-learn.
 
@@
         \section{Introduction}
+            The classification of spam emails is an important task in today's society,
+            as we receive potentially hundreds of them a year which clutter
+            an already full inbox. These emails also pose a threat to the uninformed
+            as many ask for personal information, such as street addresses and banking details.
+            Traditionally emails have been filtered through rule-based systems that
+            could be reverse engineered to still allow for spam email to get through.
+            Some newer filters take advantage of machine learning techniques and use
+            them in conjunction with the tradition method which leads to good results,
+            but still some emails get through. Some newer classifiers may hold the
+            key to filtering out even the most engineered emails to get past a filter.
 
+            As machine learning has had a resurgence a variety
+            of new classifier algorithms and implementations have surfaced that 
+            make it quick and easy to implement one of these systems. Along with this
+            an explosion of open-source libraries have made difficult parsing tasks
+            such as handling raw email data trivial. With this boom, public corpora
+            have been created through crowd sourcing, which allow individuals to
+            test their ideas on realistic data.
+
+            \begin{Figure}
+                \centering
+                \includegraphics[width=\linewidth]{figures/process.png}
+                \captionof{figure}{Our Machine Learning Process}
+                \label{fig:process}
+            \end{Figure}
+
+            In this paper we explore the selection public datasets, data preparation,
+            feature extraction and finally the selection and tuning of classifiers.
+            We will do this following the process illustrated in Figure \ref{fig:process}.
         \section{Dataset Selection}
@@
             different types of features. There were many characteristics that were
-            on our "wishlist." Such as having the freedom of processing 
+            on our `wishlist.' Such as having the freedom of processing 
             the raw text ourselves and determining what features may be important, 
@@
             origninating from a variety of sources.
+
+            TODO: needed info?
+
         \section{Data Preparation}
+
+            TODO: methodology
+
             Once data was loaded in we started by running the data the data through
             the Python Standard Library's mail parser, which striped the mail headers
-            and put them into a map for separate analysis. It also handled multipart
-            data as files were encoded into the file in base64. Without using a mail
+            and put them into a map for separate analysis. It also handled multi-part
+            data as files were encoded into the file in base64, and allowed us to count
+            the amount of attachments. Without using a mail
             parser it would be difficult to just get the body of the text which
@@
             email, possibly because this information is hard to extract and the tools
-            weren't as easily accessable when their implementation was written. We
+            weren't as easily accessible when their implementation was written. We
             decided to strip this data out as to not add noise and also because many 
@@
             it believed the message was spam or ham.
+
+            Next we used BeautifulSoup's HTML parser to interpret any HTML data that
+            was in an email. We queried the document structure for the \textit{href}
+            attribute because it indicates a link. Not only an \textit{a} tag can include
+            a link, tags such as an \textit{img} tag can also contain one. Potential attackers
+            may also use an \textit{a} tag to make a link look like it is going somewhere else,
+            such as having the text \textit{http://google.com} a clickable link to
+            \textit{http://mygooglephishingsite.com}, but we didn't see this practice 
+            occurring in this dataset so we ignored this potential feature. We also looked at the
+            amount of linked pictures, as an attacker may try to get credibility with it's prey
+            by introducing fancy graphics. Lastly we extracted the raw content from the HTML,
+            ignoring any structure that was present for tokenization.
+
+            The tokenization step was done with NLTK's word and sentence tokenizer.
+            The sentence tokens were used for some metrics such as average words per
+            sentence and could be used in the future for the analysis of sentence
+            structure. The word tokens were stripped of punctuation such as '.',
+            '?', '(' and ')', which would litter our unigrams. Next we removed stopwords
+            from the provided list of English stopwords in NLTK's corpus. Removing
+            these words are necessary as they provide no direct meaning when it comes
+            to frequency analysis of the documents and would just provide more noise
+            in the classifier. Lastly we
+            stemmed the words using NLTK's Porter Stemmer with their extensions to
+            Martin Porter's algorithm. This allowed use to remove suffixes such as '-s',
+            which would group `business' and `businesses' separately.
+
+            Now data was stored in memory as Python objects for processing through the
+            feature extractor.
+
         \section{Feature Extraction}
 
-        \section{Classifier Selection}
+        %\section{Classifier Selection}
 
-        \section{Classifier Tuning}
+        %\section{Classifier Tuning}
 
-        \section{Results}
+        %\section{Results}
 
-        \section{Evaluation}
+        %\section{Evaluation}
 
-        \section{Conclusion}
+        %\section{Conclusion}
 
-        \section{References}
-
+        %\section{References}
+        % \printbibliography
     \end{multicols}
-    %\bibliographystyle{apacite}
-    %\bibliography{spam.bib}
 \end{document}

### 2018-12-09 paper/paper.tex
Add more figures
Finished data preperation and dataset selection

@@
 \usepackage{wrapfig}
+\usepackage{subfig}
+%\usepackage{subcaption}
+
 %\usepackage{biblatex}
 \usepackage[margin=0.5in]{geometry} % don't know what size margins to use
-
 \newenvironment{Figure}
@@
             Lastly, we would like a dataset that was referenced in multiple papers to make comparisons
-            of our results to others. So we shortlisted a few datasets that fulfilled
-            some of these characteristics.
+            of our results to others.
 
-            The UCI SMS Spam Dataset is a collection of conversational text messages
-            origninating from a variety of sources.
-
-            TODO: needed info?
+            So we shortlisted a few datasets that fulfilled some of these characteristics, but
+            settled on the SpamAssasin Dataset which fulfilled all of our requirements
+            with raw email data that mostly consisted of both business and casual
+            conversations. It consisted of both easy and hard data, but we stuck with
+            just the easy data as many of the hard pieces of data were too difficult
+            to classify without more information about the user.
 
@@
 
-            TODO: methodology
+            With so much metadata given with this dataset, we wanted to take advantage
+            of it for use as potential features. In preprocessing we found many features
+            that could signature of spam emails such as excessive links or attachments.
+            The processes illustrated with Figure \ref{fig:preprocess}, is an overview
+            of the processes that will be preformed in this step.
+
+            \begin{Figure}
+                \centering
+                \includegraphics[width=\linewidth]{figures/prep.png}
+                \captionof{figure}{Preparation Process}
+                \label{fig:preprocess}
+            \end{Figure}
 
@@
             Now data was stored in memory as Python objects for processing through the
-            feature extractor.
+            feature extractor. We shuffled and then split the data 
+            80:20 into a training and testing set respectively.
 
@@
 
+            The next step is to extract the necessary features and then vectorize
+            them for training with our classifier. The first step of extraction was
+            to
+
+            \begin{minipage}{0.4\columnwidth}
+                \includegraphics[width=\textwidth]{figures/spam_wc}
+                \captionof{figure}{Spam}
+                \label{spam_wc}
+            \end{minipage}
+            \begin{minipage}{0.4\columnwidth}
+                \includegraphics[width=\textwidth]{figures/ham_wc}
+                \captionof{figure}{Ham}
+                \label{ham_wc}
+             \end{minipage}
+            This is more text
         %\section{Classifier Selection}
@@
         %\section{Classifier Tuning}
-
+            % Grid Search and Cross Validation
+            % play with Max amount of features to see if have effect
         %\section{Results}
-
+            % important metrics, why
         %\section{Evaluation}

### 2018-12-10 paper/paper.tex
add extract figure, fix references

@@
 \usepackage{subfig}
-%\usepackage{subcaption}
 
-%\usepackage{biblatex}
 \usepackage[margin=0.5in]{geometry} % don't know what size margins to use
@@
 \setlength{\columnsep}{1cm}
-%\addbibresource{paper.bib}
-%\nocite{*}
+
 \sectionfont{\fontsize{12}{15}\selectfont}
@@
                 \includegraphics[width=\linewidth]{figures/process.png}
-                \captionof{figure}{Our Machine Learning Process}
-                \label{fig:process}
+                \captionof{figure}{Our Machine Learning Process}\label{fig:process}
             \end{Figure}
@@
             with raw email data that mostly consisted of both business and casual
-            conversations. It consisted of both easy and hard data, but we stuck with
+            conversations\cite{sa_dataset}. It consisted of both easy and hard data, but we stuck with
             just the easy data as many of the hard pieces of data were too difficult
@@
                 \includegraphics[width=\linewidth]{figures/prep.png}
-                \captionof{figure}{Preparation Process}
-                \label{fig:preprocess}
+                \captionof{figure}{Preparation Process}\label{fig:preprocess}
             \end{Figure}
@@
             The next step is to extract the necessary features and then vectorize
-            them for training with our classifier. The first step of extraction was
-            to
+            them for training with our classifier.
+
+            \begin{Figure}
+                \centering
+                \includegraphics[scale=0.5]{figures/extract.png}
+                \captionof{figure}{Extraction Process}
+                \label{fig:extrac}
+            \end{Figure}
+
+            The first step of extraction was
+            to find words that meant the most to the corpus which we defined as
+            the body of the email. To accomplish this
+            we used the tf-idf algorithm implemented in scikit-learn. This algorithm helps
+            weigh down words that are too common, and bring up words that are important
+            to the corpus. Next we split the documents into spam and ham before analysis.
+            This was done because this algorithm is still based on frequency and the
+            ratio of ham to spam was not 1:1. Characteristic words to spam emails such
+            as ``free'' wouldn't show up because they're not prominent in ham emails.
 
@@
                 \includegraphics[width=\textwidth]{figures/spam_wc}
-                \captionof{figure}{Spam}
-                \label{spam_wc}
+                \captionof{figure}{Spam}\label{fig:spam_wc}
             \end{minipage}
@@
                 \includegraphics[width=\textwidth]{figures/ham_wc}
-                \captionof{figure}{Ham}
-                \label{ham_wc}
+                \captionof{figure}{Ham}\label{fig:ham_wc}
              \end{minipage}
-            This is more text
+
+             Figures \ref{fig:spam_wc} and \ref{fig:ham_wc} show
+             the most prominent 50 words in the corpus. Most the the words in the
+             ham word cloud relate to business operations asking someone to do about
+             task or getting times and dates of events. When you look at the spam
+             word cloud you see a few similar words such as ``mail'' or ``get''.
+             But in general you see words that you would associate with someone trying
+             to get you to send them money or buy a product. Words like ``money'',
+             ``offer'' and ``order'' are the epitome of this.
         %\section{Classifier Selection}
@@
 
-        %\section{Conclusion}
+        \section{Conclusion}
 
-        %\section{References}
-        % \printbibliography
+        \nocite{*}
+        \bibliography{paper} 
+        \bibliographystyle{alpha}
     \end{multicols}

### 2018-12-10 paper/paper.tex
Added the start of a conclusion

@@
              ``offer'' and ``order'' are the epitome of this.
+
+             We followed this same process for extracting features for the title, but
+             we kept the features separate as a title's content may have more distinct
+             patterns compared to the body of an email
+
         %\section{Classifier Selection}
@@
         \section{Conclusion}
-
+            This paper has shown that the usage of machine learning alone can preform
+            the classification well, compared to other's results our metrics are lower
+            than the other paper's we are comparing ourselves against. Adding more features
+            that aren't word based may help even more with performance and also increasing the
+            size of the dataset would help as well. We also recognize the limitations of
+            this implementation lends itself to only classify emails from the early 2000s, as
+            we did no testing on modern email data which may have a distribution that doesn't
+            work well with our implementation.
         \nocite{*}

### 2018-12-10 paper/paper.tex
remove typo
Add sections on classifiers, tuning, and results
Also set margin to 1" and put in autoref

--- 2018-12-09 paper/paper.tex
@@
 \usepackage{subfig}
+\usepackage{hyperref}
 %\usepackage{subcaption}
@@
 %\usepackage{biblatex}
-\usepackage[margin=0.5in]{geometry} % don't know what size margins to use
+\usepackage[margin=1.0in]{geometry} % don't know what size margins to use
 \newenvironment{Figure}
@@
 \sectionfont{\fontsize{12}{15}\selectfont}
+\subsectionfont{\fontsize{12}{15}\selectfont}
 
@@
             feature extraction and finally the selection and tuning of classifiers.
-            We will do this following the process illustrated in Figure \ref{fig:process}.
+            We will do this following the process illustrated in \autoref{fig:process}.
         \section{Dataset Selection}
@@
             that could signature of spam emails such as excessive links or attachments.
-            The processes illustrated with Figure \ref{fig:preprocess}, is an overview
+            The processes illustrated with \autoref{fig:preprocess}, is an overview
             of the processes that will be preformed in this step.
@@
             This is more text
-        %\section{Classifier Selection}
 
-        %\section{Classifier Tuning}
-            % Grid Search and Cross Validation
-            % play with Max amount of features to see if have effect
-        %\section{Results}
-            % important metrics, why
-        %\section{Evaluation}
+        \section{Classifier Selection}
+
+            We selected various classifiers from the scikit-learn Python library. The 
+            classifiers chosen for evaluation were Random Forest, K-Nearest Neighbors (KNN), 
+            Support Vector Machine (SVM) with the radial basis function (RBF) kernel, and Bernoulli 
+            Naive Bayes. All were chosen because they are well-known supervised machine learning 
+            algorithms and generally good for classification tasks.
+
+        \section{Classifier Tuning}
+            Once the classifiers were selected, we tuned them using a grid search with 5-fold 
+            cross-validation maximizing accuracy. The parameters evaluated for each classifier along 
+            with the chosen classifier parameters are as follows:
+
+            \noindent\textbf{Random Forest}
+            \begin{itemize}
+                \setlength\itemsep{0em}
+                \item Number of Estimators: 40
+                \item Maximum Tree Depth: 40
+            \end{itemize}
+            \textbf{K-Nearest Neighbors}
+            \begin{itemize}
+                \setlength\itemsep{0em}
+                \item Number of Neighbors: 10
+                \item Weight by Distance?: Yes
+            \end{itemize}
+            \textbf{SVM}
+            \begin{itemize}
+                \setlength\itemsep{0em}
+                \item Error Penalty: 10
+                \item RBF Kernel Coefficient: 0.001
+            \end{itemize}
+            \textbf{Bernoulli Naive Bayes}
+            \begin{itemize}
+                \setlength\itemsep{0em}
+                \item Learn Prior Probabilities?: No
+                \item Smoothing?: None
+            \end{itemize}
+
+            We also examined the effect of tuning the number of feature words chosen by the TF-IDF 
+            extractor. The number of feature words chosen for initial analysis was 50, but subsequent 
+            tests ranged from 50 to 1050 in steps of 200.
+        \section{Results}
+            Classifiers were evaluated on precision and recall on the spam class. Overall, precision is
+            weighted more heavily than recall in the evaluation because it would likely be more disruptive 
+            to classify a ham email as spam than to let a spam message slip through.
+            \subsection{General Precision and Recall}
+                Overall, the classifiers performed well with just features extracted from the email body.
+                Precision and recall comparison can be found in \autoref{fig:precision_recall} with direct 
+                number comparison in \autoref{tab:precision_recall}.
+
+                \begin{Figure}
+                    \begin{tabular}{|l|l|l|}
+                    \hline
+                                  & Precision & Recall \\ \hline
+                    KNN           & 90.91     & 51.02  \\ \hline
+                    Naive Bayes   & 83.67     & 83.67  \\ \hline
+                    Random Forest & 98.84     & 86.73  \\ \hline
+                    SVM           & 83.33     & 40.82  \\ \hline
+                    \end{tabular}
+                    \captionof{table}{Precison and Recall}
+                    \label{tab:precision_recall}
+                \end{Figure}
+
+                \begin{Figure}
+                    \centering
+                    \includegraphics[width=\linewidth]{figures/precision_recall.png}
+                    \captionof{figure}{Precison and Recall Comparison}
+                    \label{fig:precision_recall}
+                \end{Figure}
+
+                While all classifiers had a precision greater than 80\%, the recall clearly divides the 
+                classifiers into two groups. Those with over 80\% recall were Random Forest and Naive Bayes, 
+                while the KNN and SVM both had a recall lower than 50\%.
+
+                Overall, the Random Forest classifier performed the best under our evaluation criteria 
+                with an almost perfect precision. In fact, only one ham instance was misclassified as spam. 
+                Random Forest also performed best on recall. The Naive Bayes classifier performed second best 
+                with an even recall and precision.
+
+            \subsection{Effect of Number of Feature Words} 
+                When increasing the number of feature words, precision increased dramatically for some classifiers. 
+                Specifically, Naive Bayes and the SVM gained a precision increase of 10\% and 8\% respectively (See \autoref{fig:precision_word_effect}).
+                However, the number of feature words did not seem to effect recall as much as precision (See \autoref{fig:recall_word_effect}).
+                Only one classifier gained over 3\% precision from 50 feature words. Naive Bayes gained 10\% recall when increasing 
+                the number of feature words from 50 to 250. This gain dropped off as the number of feature words increased.
+
+                \begin{Figure}
+                    \centering
+                    \includegraphics[width=\linewidth]{figures/recall_wv_effect.png}
+                    \captionof{figure}{Effect of the number of feature words on classifier recall}
+                    \label{fig:recall_word_effect}
+                \end{Figure}
+
+                \begin{Figure}
+                    \centering
+                    \includegraphics[width=\linewidth]{figures/precision_wv_effect.png}
+                    \captionof{figure}{Effect of the number of feature words on classifier precision}
+                    \label{fig:precision_word_effect}
+                \end{Figure}
+
+                Interestingly, Naive Bayes dropped in precision by over 2\% 
+                for 250 and 450 feature words before having the largest gain of any classifier. 
+                Possibly, the gain in recall had a negative effect on the precision.
 

### 2018-12-10 paper/paper.tex

@@
     \begin{center}
-        \textbf{Evaluation of Classifiers and Feature Selection on Spam Filtering test} 
+        \textbf{Evaluation of Classifiers and Feature Selection on Spam Filtering} 
     \end{center}
@@
                 \includegraphics[width=\textwidth]{figures/ham_wc}
-                \captionof{figure}{Ham}
-                \label{ham_wc}
+                \captionof{figure}{Ham}\label{fig:ham_wc}
              \end{minipage}
-            This is more text
+
+             Figures \ref{fig:spam_wc} and \ref{fig:ham_wc} show
+             the most prominent 50 words in the corpus. Most the the words in the
+             ham word cloud relate to business operations asking someone to do about
+             task or getting times and dates of events. When you look at the spam
+             word cloud you see a few similar words such as ``mail'' or ``get''.
+             But in general you see words that you would associate with someone trying
+             to get you to send them money or buy a product. Words like ``money'',
+             ``offer'' and ``order'' are the epitome of this.
+
+             We followed this same process for extracting features for the title, but
+             we kept the features separate as a title's content may have more distinct
+             patterns compared to the body of an email
 
@@
 
+        %\section{Classifier Selection}
+
                 \begin{Figure}
@@
         % \printbibliography
+        \section{Conclusion}
+
+            This paper has shown that the usage of machine learning alone can preform
+            the classification well, compared to other's results our metrics are lower
+            than the other paper's we are comparing ourselves against. Adding more features
+            that aren't word based may help even more with performance and also increasing the
+            size of the dataset would help as well. We also recognize the limitations of
+            this implementation lends itself to only classify emails from the early 2000s, as
+            we did no testing on modern email data which may have a distribution that doesn't
+            work well with our implementation.
+        \nocite{*}
+        \bibliography{paper} 
+        \bibliographystyle{alpha}
     \end{multicols}

### 2018-12-10 paper/paper.tex
Add a couple paragraphs. Some small corrections.

@@
             weighted more heavily than recall in the evaluation because it would likely be more disruptive 
-            to classify a ham email as spam than to let a spam message slip through.
+            to classify a ham email as spam than to let in an occasional spam message. Subsequently, the 
+            classifiers were also subjected to various feature word counts to observe the effect.
             \subsection{General Precision and Recall}
@@
                         SVM                                 & 83.33     & 40.82  \\ \hline
+                        Naive Bayes \cite{hovold2005naive}  & 98.46     & 99.66  \\ \hline
+                        DBM \cite{tzortzis2007deep}         & 96.40     & 95.51  \\ \hline
+                        SVM \cite{tzortzis2007deep}         & 96.14     & 95.24  \\ \hline
                     \end{tabular}
@@
 
+                Comparing with classifiers by other researchers in \autoref{tab:precision_recall}, the 
+                Random Forest classifier performs better in precision than all three classifiers by other researchers. 
+                Looking at recall, however, leaves much to be desired. Likely due to focusing solely on email body 
+                content for text analysis, our implementation misses important information in subject lines or 
+                email addresses.
+
             \subsection{Effect of Number of Feature Words} 
@@
 
-        %\section{Classifier Selection}
-
                 \begin{Figure}
@@
 
-        %\section{Conclusion}
-
-        %\section{References}
-        % \printbibliography
         \section{Conclusion}
 
-            This paper has shown that the usage of machine learning alone can preform
-            the classification well, compared to other's results our metrics are lower
-            than the other paper's we are comparing ourselves against. Adding more features
+            This paper has shown that the usage of machine learning alone on email body text 
+            along with a few other features can preform the classification well. 
+            However, our metrics are lower
+            than the other papers we are comparing against ourselves. Adding more features
             that aren't word based may help even more with performance and also increasing the
             size of the dataset would help as well. We also recognize the limitations of
-            this implementation lends itself to only classify emails from the early 2000s, as
+            this implementation lending itself to only classify emails from the early 2000s, as
             we did no testing on modern email data which may have a distribution that doesn't

### 2018-12-10 paper/paper.tex
Spelling grammar corrections

@@
             This paper has shown that the usage of machine learning alone on email body text 
-            along with a few other features can preform the classification well. 
+            along with a few other features can perform the spam classification task well. 
             However, our metrics are lower

### 2018-12-10 paper/paper.tex
Grammar correction

@@
             Some newer filters take advantage of machine learning techniques and use
-            them in conjunction with the tradition method which leads to good results,
+            them in conjunction with the traditional method which leads to good results,
             but still some emails get through. Some newer classifiers may hold the
-            key to filtering out even the most engineered emails to get past a filter.
+            key to filtering out even the most engineered emails.
 
-            As machine learning has had a resurgence a variety
+            As machine learning has had a resurgence, a variety
             of new classifier algorithms and implementations have surfaced that 
-            make it quick and easy to implement one of these systems. Along with this
+            make it quick and easy to implement one of these systems. Along with this,
             an explosion of open-source libraries have made difficult parsing tasks

### 2018-12-14 spam_classifier/classifier.py
Refactoring, adding changing classifier to CLI, documentation, added title features and capital frequency features

@@ -1,12 +1,13 @@
-from sklearn.naive_bayes import BernoulliNB
-from sklearn.ensemble import RandomForestClassifier
-from sklearn.neighbors import KNeighborsClassifier
+"""
+Wrapper class for training model
+
+"""
+
 from sklearn.metrics import confusion_matrix, classification_report
 import numpy
 
-from sklearn import svm, model_selection, neighbors
+from sklearn import model_selection, neighbors
 from .plot import plot_confusion_matrix 
-
 
 
 class StatelessClassifier:
@@ -16,72 +17,57 @@ class StatelessClassifier:
 
     """
 
-    def train(self, train_data, train_label):
-        print("Training...")
+    __slots__ = ['train_data', 'train_label']
 
-        train_data_matrix = numpy.asarray(train_data, dtype=object)
+    def __init__(self, train_data, train_label):
+        self.train_data = train_data
+        self.train_label = train_label
+
+    def train(self, classifier, grid_search_params=None, train_folds=5):
+        """
+        Train a classifier with the given data.
+
+        :param train_data training documents
+        :param train_label training labels that correspond to data train_data[i] = train_label[i]
+
+        :return model with the best esimator
+        """
+
+        print("Training...")
+        train_data_matrix = numpy.asarray(self.train_data, dtype=object)
         print("Train data shape: ", train_data_matrix.shape)
 
-        train_label_matrix = numpy.asarray(train_label, dtype=object)
+        train_label_matrix = numpy.asarray(self.train_label, dtype=object)
         print("Train label shape: ", train_label_matrix.shape)
 
-        # grid_search = self.grid_search_svc(train_data_matrix, train_label_matrix, 5)
-        # grid_search = self.grid_search_knn(train_data_matrix, train_label_matrix, 5)
-        # grid_search = self.grid_search_nb(train_data_matrix, train_label_matrix, 5)
-        grid_search = self.grid_search_rf(train_data_matrix, train_label_matrix, 5)
+        cla = None
 
-        print(grid_search.best_params_)
-        # print(grid_search.best_estimator_.feature_importances_)
-        return grid_search.best_estimator_
+        if grid_search_params != None:
+            classifier = self.grid_search(classifier, grid_search_params, train_folds)
+            classifier.fit(self.train_data, self.train_label)
[…]

### 2018-12-14 spam_classifier/cli.py

@@ -1,3 +1,8 @@
+"""
+CLI for the classifier.
+
+"""
+
 import os
 
 import click
@@ -21,8 +26,9 @@ def data(data_dir, output_dir):
 @cli.command()
 @click.argument("training_file", type=click.Path(exists=True, dir_okay=False, file_okay=True))
 @click.argument("classifier_out", type=click.Path(dir_okay=False, file_okay=True))
-def train(training_file, classifier_out):
-    main.train(training_file, classifier_out)
+@click.argument("classifier_type", type=click.Choice(['svc', 'knn', 'nb', 'rf']))
+def train(training_file, classifier_out, classifier_type):
+    main.train(training_file, classifier_out, classifier_type)
 
 
 @cli.command()
@@ -36,11 +42,12 @@ def test(test_file, classifier_file, errors_out):
 @cli.command()
 @click.argument("data_dir", type=click.Path(dir_okay=True, file_okay=False))
 @click.argument("output_dir", type=click.Path(dir_okay=True, file_okay=False))
-def auto(data_dir, output_dir):
+@click.argument("classifier_type", type=click.Choice(['svc', 'knn', 'nb', 'rf']))
+def auto(data_dir, output_dir, classifier_type):
     classifier_out = os.path.join(output_dir, "classifier.pkl")
     errors_out = os.path.join(output_dir, "errors.csv")
     train_out, test_out = main.data(data_dir, output_dir)
-    main.train(train_out, classifier_out)
+    main.train(train_out, classifier_out, classifier_type)
     main.test(test_out, classifier_out, errors_out)
 
 

### 2018-12-14 spam_classifier/features.py

@@ -1,3 +1,8 @@
+"""
+Extractor of text features.
+
+"""
+
 import re
 import string
 from collections import Counter
@@ -12,6 +17,10 @@ from bs4 import BeautifulSoup
 
 
 class LambDocument(object):
+    """
+    A LambDocument is a document that hasn't been classified, and to be used to
+    extract features for classification.
+    """
 
     __slots__ = [
         'idd',
@@ -20,7 +29,9 @@ class LambDocument(object):
         'label',
         'sents',
         'tokens',
+        'title_tokens',
         'word_freq',
+        'title_word_freq',
         'num_attach',
         'num_word',
         'avg_word_len'
@@ -39,35 +50,87 @@ class LambDocument(object):
         self.num_word = len(raw_text)
 
         # Preprocess text
-        raw_text = raw_text.lower() # all lower case
         trans = str.maketrans('', '', string.punctuation)
 
-        raw_text = raw_text.translate(trans) # strip puncuation
+        raw_text = raw_text.lower().translate(trans) # strip puncuation
         self.tokens = nltk.tokenize.word_tokenize(raw_text, language='english')
+
+        title = mail.subject.lower().translate(trans)
+        self.title_tokens = nltk.tokenize.word_tokenize(title, language='english')
+        
         self.avg_word_len = reduce(lambda x, y: x + len(y), self.tokens, 0) / (len(self.tokens) + 1)
+        
         self.tokens = [w for w in self.tokens if not w in stopwords.words('english')]
         stemmer = PorterStemmer()
         self.tokens = [stemmer.stem(w) for w in self.tokens] # remove stems
 
+        self.title_tokens = [w for w in self.title_tokens if not w in stopwords.words('english')]
+        self.title_tokens = [stemmer.stem(w) for w in self.title_tokens] # remove stems
+
         self.word_freq = Counter(self.tokens)
+        self.title_word_freq = Counter(self.title_tokens)
 
     def num_sent(self):
-        """ Get the total number of sentences. """
+        """ 
+        Get the total number of sentences. 
+        
+        :return number of sentences
+        """
         return len(self.sents)
 
     def avg_sen_len(self):
+        """
+        Get the average sentence length.
+
+        :return average sentence length
+        """
         return self.num_word // (1+self.num_sent())
 
     def count_links(self):
+        """
+        Count the amount of links based on HTML href.
+
+        :return amount of links
+        """
         return len(self.text_soup.find_all(href=True))
 
     def count_images(self):
[…]

### 2018-12-14 spam_classifier/main.py

@@ -1,21 +1,62 @@
+"""
+Implementation of main functions of the classifier.
+
+"""
+
 import os
 import csv
 from sklearn.externals import joblib
 from spam_classifier import corpus_reader, classifier
 
+from sklearn.naive_bayes import BernoulliNB
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.svm import SVC
+
+rf_param_grid = {
+    'n_estimators': [5, 10, 20, 40], 
+    'max_depth': [5, 10, 20, 40, None], 
+    'max_features': ['auto', 'log2', None]
+}
+
+svc_param_grid = {
+    'C': [0.001, 0.01, 0.1, 1, 10], 
+    'gamma': [0.001, 0.01, 0.1, 1]
+}
+
+nb_param_grid = {
+    'fit_prior': [True, False], 
+    'alpha': [1.0e-10, 1.0]
+}
+
+knn_param_grid = {
+    'n_neighbors': [1, 3, 5, 10, 15, 20], 
+    'weights': ['uniform', 'distance']
+}
 
 def data(data_dir, output_dir):
+    """
+    Load data, create vectors and save to output directory.
+
+    :param data_dir directory to save data
+    :param output_dir directory to output vectors
+
+    :return train_path name of file the training data saved too
+    :return test_path name of file the test data saved too
+    """
+
     print("Loading in data...")
     d = corpus_reader.load_data(data_dir=data_dir)
 
     train, test = corpus_reader.split_documents(d)
 
     print("Making vectors...")
-    features = corpus_reader.find_feature_words(train, 50)
-    train_v = corpus_reader.get_vectors(train, features)
-    test_v = corpus_reader.get_vectors(test, features)
+    body_features = corpus_reader.find_feature_words(train, num_feature=50)
+    title_features = corpus_reader.find_feature_words(train, num_feature=50, token_parameter='title_word_freq')
+    train_v = corpus_reader.get_vectors(train, body_features, title_features)
+    test_v = corpus_reader.get_vectors(test, body_features, title_features)
 
-    data_description = train[0].get_vector_description(features)
+    data_description = train[0].get_vector_description(body_features, title_features)
 
     os.makedirs(output_dir, exist_ok=True)
     train_path = os.path.join(output_dir, "train.csv")
@@ -30,23 +71,56 @@ def data(data_dir, output_dir):
     return train_path, test_path
 
 
-def train(training_file, classifier_out):
[…]

### 2018-12-14 spam_classifier/plot.py

@@ -1,3 +1,8 @@
+"""
+Graphing functions for evaluation.
+
+"""
+
 import itertools
 from collections import Counter
 
@@ -8,6 +13,16 @@ import nltk.tokenize
 
 # https://scikit-learn.org/stable/auto_examples/model_selection/plot_confusion_matrix.html
 def plot_confusion_matrix(cm, classes, normalize=False, title='Confusion matrix', filename='conf_matrix.png', cmap=plt.cm.Reds):
+    """
+    Plot a confusion matrix with matplotlib.
+
+    :cm confusion matrix
+    :classes classes of graph as list
+    :normalize normalize to 0-1 scale
+    :title title of graph
+    :filename filename to save too
+    :cmap color map to use
+    """
     if normalize:
         cm = cm.astype('float') / cm.sum(axis=1)[:, np.newaxis]
         print("Normalized confusion matrix")
@@ -36,6 +51,14 @@ def plot_confusion_matrix(cm, classes, normalize=False, title='Confusion matrix'
     plt.savefig(filename)
 
 def plot_wordcloud(freq, corpus, filename):
+    """
+    Create a wordcloud from the frequently used words.
+
+    :freq list of words to extract
+    :corpus list of all documents
+    :filename name of file to save
+    """
+
     tokens = []
     
     for doc in corpus:

### 2018-12-14 spam_classifier/classifier.py
Fixed CLI

@@ -45,7 +45,7 @@ class StatelessClassifier:
         if grid_search_params != None:
             classifier = self.grid_search(classifier, grid_search_params, train_folds)
             classifier.fit(self.train_data, self.train_label)
-            cla = classifier.best_estimator
+            cla = classifier.best_estimator_
         else:
             cla = classifier.fit(self.train_data, self.train_label)
 

### 2018-12-14 spam_classifier/features.py

@@ -144,7 +144,7 @@ class LambDocument(object):
             total_capital
         ]
 
-        vec += per_len_capital
+        vec += [x[1] for x in per_len_capital]
         vec += self.title_word_freq_select(title_word_select_features)
         vec += self.word_freq_select(word_select_features)
 

### 2018-12-14 spam_classifier/main.py

@@ -126,7 +126,7 @@ def test(test_file, classifier_file, errors_out):
     test_file_names = [x[0] for x in test_csv]
     test_data = [x[2:] for x in test_csv]
 
-    use_ids = classifier.StatelessClassifier()
+    use_ids = classifier.StatelessClassifier(None, None)
 
     final_classifier = joblib.load(filename=classifier_file)
 

### 2018-12-14 paper/paper.tex
Remove section about subject lines until we actually do it
Update images, add more detail to report, update diagrams to fit

@@
 \usepackage{hyperref}
-%\usepackage{subcaption}
+\usepackage[margin=1.0in]{geometry} % don't know what size margins to use
 
-%\usepackage{biblatex}
-\usepackage[margin=1.0in]{geometry} % don't know what size margins to use
 \newenvironment{Figure}
@@
         of other models introduced by other authors. Classifiers will be selected
-        from the variety available through the open-source Python module scikit-learn.
+        from the variety available through the open-source Python module scikit-learn \cite{scikit-learn}.
 
-    \begin{multicols}{2}
+    \begin{multicols*}{2}
         \section{Introduction}
@@
                 \centering
-                \includegraphics[width=\linewidth]{figures/process.png}
+                \includegraphics[width=0.5\linewidth]{figures/process.png}
                 \captionof{figure}{Our Machine Learning Process}\label{fig:process}
@@
             So we shortlisted a few datasets that fulfilled some of these characteristics, but
-            settled on the SpamAssasin Dataset which fulfilled all of our requirements
+            settled on the SpamAssassin Dataset which fulfilled all of our requirements
             with raw email data that mostly consisted of both business and casual
@@
             Next we used BeautifulSoup's HTML parser to interpret any HTML data that
-            was in an email. We queried the document structure for the \textit{href}
+            was in an email \cite{beautiful_soup}. We queried the document structure for the \textit{href}
             attribute because it indicates a link. Not only an \textit{a} tag can include
@@
             sentence and could be used in the future for the analysis of sentence
-            structure. The word tokens were stripped of punctuation such as '.',
+            structure \cite{nltk}. The word tokens were stripped of punctuation such as '.',
             '?', '(' and ')', which would litter our unigrams. Next we removed stopwords
@@
             stemmed the words using NLTK's Porter Stemmer with their extensions to
-            Martin Porter's algorithm. This allowed use to remove suffixes such as '-s',
+            Martin Porter's algorithm \cite{porter1980algorithm}. This allowed use to remove suffixes such as '-s',
             which would group `business' and `businesses' separately.
@@
             The next step is to extract the necessary features and then vectorize
-            them for training with our classifier.
+            them for training with our classifier. We used a combination of features
+            extracted from the structure of the email and text, along with traditional
+            word frequency for important words.
 
@@
                 \includegraphics[scale=0.5]{figures/extract.png}
-                \captionof{figure}{Extraction Process}
-                \label{fig:extrac}
+                \captionof{figure}{Extraction Process}\label{fig:extrac}
             \end{Figure}
@@
 
-            \begin{minipage}{0.4\columnwidth}
+            \begin{minipage}{0.48\columnwidth}
                 \includegraphics[width=\textwidth]{figures/spam_wc}
@@
             \end{minipage}
-            \begin{minipage}{0.4\columnwidth}
+            \begin{minipage}{0.48\columnwidth}
                 \includegraphics[width=\textwidth]{figures/ham_wc}
@@
              we kept the features separate as a title's content may have more distinct
-             patterns compared to the body of an email
+             patterns compared to the body of an email.
+
+             We also added some analysis of sequences of capital letters, as a characteristic
+             of spam emails is all caps sentences to grab the attention of the receiver. We
+             kept track of sequences that range in length from 3 to 10, individually recording
+             the frequency of occurrence.
 
@@
         \bibliographystyle{alpha}
-    \end{multicols}
+    \end{multicols*}
 \end{document}

### 2018-12-16 paper/paper.tex
Formatting fixes. Make line graphs pareable in B&W
Small grammar corrections

@@
 
-    \begin{multicols*}{2}
+    \begin{multicols}{2}
         \section{Introduction}
@@
             occurring in this dataset so we ignored this potential feature. We also looked at the
-            amount of linked pictures, as an attacker may try to get credibility with it's prey
+            amount of linked pictures, as an attacker may try to get credibility with their prey
             by introducing fancy graphics. Lastly we extracted the raw content from the HTML,
-            ignoring any structure that was present for tokenization.
+            ignoring any structure that was present to prepare for tokenization.
 
@@
             to frequency analysis of the documents and would just provide more noise
-            in the classifier. Lastly we
+            in the classifier. Lastly, we
             stemmed the words using NLTK's Porter Stemmer with their extensions to
@@
                 \centering
-                \includegraphics[scale=0.5]{figures/extract.png}
+                \includegraphics[width=\linewidth]{figures/extract.png}
                 \captionof{figure}{Extraction Process}\label{fig:extrac}
@@
             ratio of ham to spam was not 1:1. Characteristic words to spam emails such
-            as ``free'' wouldn't show up because they're not prominent in ham emails.
-
-            \begin{minipage}{0.48\columnwidth}
+            as ``free'' wouldn't show up because they're not prominent in ham emails.\\\\
+            \noindent\begin{minipage}{0.48\columnwidth}
                 \includegraphics[width=\textwidth]{figures/spam_wc}
@@
              Figures \ref{fig:spam_wc} and \ref{fig:ham_wc} show
-             the most prominent 50 words in the corpus. Most the the words in the
+             the most prominent 50 words in the corpus. Most of the words in the
              ham word cloud relate to business operations asking someone to do about
@@
             tests ranged from 50 to 1050 in steps of 200.
+
             \section{Results}
@@
             classifiers were also subjected to various feature word counts to observe the effect.
+    \end{multicols}
+        \begin{Figure}
+            \centering
+            \includegraphics[width=\linewidth]{figures/precision_recall.png}
+            \captionof{figure}{Precison and Recall Comparison}
+            \label{fig:precision_recall}
+        \end{Figure}
+    \begin{multicols}{2}
             \subsection{General Precision and Recall}
@@
 
-                \begin{Figure}
-                    \centering
-                    \includegraphics[width=\linewidth]{figures/precision_recall.png}
-                    \captionof{figure}{Precison and Recall Comparison}
-                    \label{fig:precision_recall}
-                \end{Figure}
-
                 While all classifiers had a precision greater than 80\%, the recall clearly divides the 
@@
 
-            \subsection{Effect of Number of Feature Words} 
-                When increasing the number of feature words, precision increased dramatically for some classifiers. 
-                Specifically, Naive Bayes and the SVM gained a precision increase of 10\% and 8\% respectively (See \autoref{fig:precision_word_effect}).
-                However, the number of feature words did not seem to effect recall as much as precision (See \autoref{fig:recall_word_effect}).
-                Only one classifier gained over 3\% precision from 50 feature words. Naive Bayes gained 10\% recall when increasing 
-                the number of feature words from 50 to 250. This gain dropped off as the number of feature words increased.
-
+    \end{multicols}
         \begin{Figure}
@@
         \end{Figure}
+    \begin{multicols}{2}
+            \subsection{Effect of Number of Feature Words} 
+                When increasing the number of feature words, precision increased dramatically for some classifiers. 
+                Specifically, Naive Bayes and the SVM gained a precision increase of 10\% and 8\% respectively (See \autoref{fig:precision_word_effect}).
+                However, the number of feature words did not seem to effect recall as much as precision (See \autoref{fig:recall_word_effect}).
+                Only one classifier gained over 3\% precision from 50 feature words. Naive Bayes gained 10\% recall when increasing 
+                the number of feature words from 50 to 250. This gain dropped off as the number of feature words increased.
 
@@
         \bibliographystyle{alpha}
-    \end{multicols*}
+    \end{multicols}
 \end{document}
